Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-11-30 paper/sec-results.tex
more work on paper

In this section we present the results of the 
cut-based analysis and
compare our results with other recent related studies.
%
We study how the signal significance
is modified if only the $4b$ component of the
QCD multi-jet background is taken into account,
but the $2b2j$ and $4j$ components are neglected.
%

@@

First of all, we compare the cross-sections at various
analysis levels for the three categories, for signal and background events,
by providing a detailed cut-flow.
%
We consider all backgrounds enumerated in Sect.~\ref{mcgeneration},
and also discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, for the case of the
 HL-LHC with
an integrated luminosity of $\mathcal{L}_{\rm int}
=3000$ fb$^{-1}$.
%
%
We emphasize that it is important not only to achieve a good signal
significance, $S/\sqrt{B}$, but also a good signal over background ratio $S/B$:
if the latter is too small, a very accurate
measurement of the background with an infeasible small
systematic uncertainty would be required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  Generator level &  Generator level & Generator level \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
 $N_{\rm jets}^{R04}\ge 4$ \\
{\bf  C1b} & +$p_T$ cuts & +$p_T$ cuts
 & +$p_T$ cuts \\
 {\bf C1c} & + rapidity cuts &  + rapidity cuts &
 + rapidity cuts \\
 {\bf C1d} & + Mass-Drop tag required & $N_{\rm jets}^{R}=1$ with MDT  &
  Higgs reco possible \\
{\bf C1e} & $m_H$ window cut & $m_H$ window cut & $m_H$ window cut  \\
\hline
{\bf C2} & $b$-tagging  &  $b$-tagging &  $b$-tagging \\
\hline
  \end{tabular}
  \caption{\small Details of the cut-flow of the present analysis in the boosted, intermediate
    and resolved categories.
    %
    See text for more details.
    \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To begin with,
in Table~\ref{tab:cutflowdetails}
we summarize the details of the cut-flow of the present analysis,
separated into the exclusive boosted, intermediate
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    The various cuts
    are defined as follows
    \begin{itemize}
    \item {\bf C0}: these are the generator level cross-sections, with
      mild generation cuts in the case of the background processes.
    \item {\bf C1a}: at this level, we only check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed here.
    \item {\bf C1b}: Same as C1a, but now the jets should
      satisfy the corresponding $p_T$ thresholds:
      $p_T \ge 200$ GeV for large-$R$ jets,
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: in addition, the above jets must satisfy the
      rapidity acceptance requirement.
    \item {\bf C1d}: at this level, large-$R$ jets must satisfy the
      mass-drop tag condition.
    \item {\bf C1e}: after the two Higgs candidate jets have been reconstructed,
      their invariant masses are required to lie within the
      $m_H$ window, namely
    \item {\bf C2}: finally, this is the level of the analysis
      after $b$-tagging has been performed.
      \end{itemize}
    Signal and background events that survive the C2 level cuts
    are then used as input for the MVA training, as described
    in Sect.~\ref{sec:mva}.

    %
    In Table~\ref{tab:cutflow_noPU_1} we provide
    the values of cross-sections in fb
     for the signal
      and the various background
      components as a function of the
      cut-flow, for the resolved,
      boosted and intermediate
      categories.
      %
      We see that although of course
      $b$-tagging suppressed very substantially the $2b2j$ and
      $4j$ backgrounds, it does not completely
      eliminate them.

\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_noPU_3.tex}
    \caption{\small The cross-sections, in pb,
      for the signal and the various background
      processes at different points of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component.
 \label{tab:cutflow_noPU_1}}
\end{table}

%
Let us begin with the discussion of the boosted category, the one that
leads to a higher signal significance.
%
We see that after all the cuts, including $b$-tagging,
we end up with almost 200 signal events at the HL-LHC, with
still a substantial background of around $40k$ events.
%
The signal significance is around 1.0 at the end of the
cut-based analysis.
%
While it should have been possible to increase this significance
but using more aggressive cuts, we have refrained doing so
since we prefer this optimisation to be performed at the level of
the MVA.

From the results of Table~\ref{tab:cutflow_noPU_1}
it is possible to compute the corresponding numbers
for the end of Run II at the LHC with
$\mathcal{L}_{\rm int}=300$ fb$^{-1}$: in the boosted category,
we have only
19 signal events, and the signal significance drops down to
$S/\sqrt{B}\sim 0.3$.
%
We will discuss in the next section what are the Run II prospects
once we include the effects of the MVA in the analysis, but
it is clear already at this level that one really needs the full
integrated luminosity of the HL-LHC to be able to observe
the Higgs pair production process, unless  of course production rates are
increased by new BSM dynamics

Both the intermediate and resolved categories benefit from higher signal yields,
specially in the resolved category, but this enhancement is washed out by the stronger
increase in the QCD multi-jet background.
%
In both cases the signal significance is around half of that in the boosted category,
with in addition $S/B$ being an order of magnitude smaller.
%
Therefore, the boosted category is clearly determined to be the most useful category,
specially due to the significant suppression of the QCD multi-jet background.
%
And we still have not exploited all the rich information contained on the jet
substructure, as we will do in the next section.

We observe
that after the cut-based analysis, the significance of the Higgs pair production
observation in the $b\bar{b}b\bar{b}$ channel using the boosted topology
is $S/\sqrt{B}=0.99$,
and that the combination
of the boosted, intermediate and resolved categories gives a slightly higher
value, around 1.20.
%

>>>>

In this section we present the results of the 
cut-based analysis.
%
We present a detailed cut-flow for the various steps
of the analysis.
%
We
compare our results with other recent related studies, in particular
we study how the signal significance
is modified if only the $4b$ component of the
QCD multi-jet background is taken into account,
but the $2b2j$ and $4j$ components are neglected.
%

@@

First of all, we compare the cross-sections at various
analysis levels for the three categories, for signal and background events,
by providing a detailed cut-flow.
%
We consider all backgrounds enumerated in Sect.~\ref{mcgeneration},
and also discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, corresponding to the case of the
 HL-LHC with
an integrated luminosity of $\mathcal{L}_{\rm int}
=3000$ fb$^{-1}$.
%
We emphasize that it is important not only to achieve a good signal
significance, $S/\sqrt{B}$, but also a good signal over background ratio $S/B$:
if the latter is too small, a very accurate
measurement of the background with an infeasible small
systematic uncertainty would be required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  \multicolumn{3}{c|}{Generator level} \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
$N_{\rm jets}^{R04}\ge 4$ \\
\hline
{\bf  C1b} & \multicolumn{3}{c|}{+$p_T$ cuts} \\
{\bf C1c} & \multicolumn{3}{c|}{+rapidity cuts}\\
\hline
 {\bf C1d} & +$N_{\rm MDT}\ge 2$ & $N_{\rm jets}^{R10}=1$ with MDT  &
 Higgs reconstruction \\
 \hline
{\bf C1e} & \multicolumn{3}{c|}{ $m_H$ window cut} \\
\hline
{\bf C2} & \multicolumn{3}{c|}{$b$-tagging}    \\
\hline
  \end{tabular}
  \caption{\small Details of the cut-flow of the present analysis in the boosted, intermediate
    and resolved categories.
    %
      \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To begin with,
in Table~\ref{tab:cutflowdetails}
we summarize the details of the cut-flow of the present analysis,
separated into the exclusive boosted, intermediate
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    The various cuts
    are defined as follows
    \begin{itemize}
    \item {\bf C0}: these are the generator level cross-sections, with
      mild generation cuts in the case of the background processes, and no cuts
      for the signal events.
      %
    \item {\bf C1a}: at this level, we only check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed here.
    \item {\bf C1b}: in addition, now the jets should
      satisfy the corresponding $p_T$ thresholds:
      $p_T \ge 200$ GeV for large-$R$ jets,
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: in addition, the above jets must satisfy the
      rapidity acceptance requirement.
    \item {\bf C1d}: at this level, the two leading large-$R$ jets must
      be mass-drop tagged in the boosted category.
      %
      In the intermediate category, the large-$R$ jet must also be mass-drop tagged.
      %
    \item {\bf C1e}: after the two Higgs candidate jets have been reconstructed,
      their invariant masses are required to lie a window around $m_H$,
      in particular between 100 and 150 GeV.
          \item {\bf C2}: This is the level of the analysis
            after $b$-tagging has been performed, using the strategy discussed
            in Sect.~\ref{sec:btagging}.
      \end{itemize}
    The signal and background events that satisfy all the analysis cut up to the
    C2 level
    are then used as input for the MVA training, as described next
    in Sect.~\ref{sec:mva}.

    %
    In Table~\ref{tab:cutflow_noPU_1} we provide
    the values of cross-sections, in femtobarns,
     for the signal
      and the various background
      components as a function of the
      cut-flow, for the resolved,
      boosted and intermediate
      categories.
      %
       In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, at the HL-LHC, considering either
      the total background or only the $4b$ component.
     %
      A first important observation from Table~\ref{tab:cutflow_noPU_1}
      is the role of the fake component of the QCD background.
      %
      We see that, even after $b$-tagging, the reducible $2b2j$ background
      is comparable or even larger, than the irreducible
      $4b$ component.
      %
      This is specially marked in the boosted category.
      %
      Therefore, at this level of the analysis, the signal significance
      can be artificially increased by up to a factor 2 if only the
      $4b$ background is considered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_noPU_3.tex}
    \caption{\small The cross-sections, in femtobarns,
      for the signal and the various background
      processes at different steps of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component.
      %
      The different levels of the cut-flow are summarized
      in Table~\ref{tab:cutflowdetails}.
 \label{tab:cutflow_noPU_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
In the boosted category, we see that after all the cuts, including $b$-tagging,
we end up with more than 1000 signal events at the HL-LHC, however swamped in a total
background of almost 1M QCD events.
%
This leads to a signal significance of around 1.0 at the end of the
cut-based analysis.
%
While it should have been possible to increase this significance
but using more aggressive cuts, we have refrained doing so
since we prefer this optimisation to be performed at the level of
the MVA.
%
We also note that $S/B$ is at the per-mille level, requiring a extremely accurate
measurement of the QCD background.

From the results of Table~\ref{tab:cutflow_noPU_1}
it is possible to compute the corresponding numbers
for the end of Run II at the LHC with
$\mathcal{L}_{\rm int}=300$ fb$^{-1}$: in the boosted category,
we have only
120 signal events, and the signal significance drops down to
$S/\sqrt{B}\sim 0.3$.
%
We will discuss in the next section what are the Run II prospects
once we include the effects of the MVA in the analysis, but
it is clear already at this level that one really needs the full
integrated luminosity of the HL-LHC to be able to observe
the Higgs pair production process, unless  of course production rates are
increased by new BSM dynamics

Both the intermediate and resolved categories benefit from higher signal yields,
specially in the resolved category, but this enhancement is compensated by the stronger
increase in the QCD multi-jet background.
%
In both cases the signal significance is similar to that of the boosted category,
with the additional drawback in the resolved case that the signal over background ratio
is one order to magnitude smaller.
%
Overall, we observe that 
that after the cut-based analysis,
the combination
of the boosted, intermediate and resolved categories yields
to total significance for the observation of the Higgs pair production
observation in the $b\bar{b}b\bar{b}$ final
state of around $S/\sqrt{B}=1.70$.
